---
title: "Nested Resource Competition Models"
author: "Ruby An"
date: "2023-02-01"
output:
  html_document:
    number_sections: true
  pdf_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)

# load required libraries 
library(deSolve)
library(seqinr)
library(tidyverse)
library(cowplot)
library(patchwork)
library(tinytex)

# set plot theme
theme_set(theme_cowplot())

# colors for plotting
color_palette <- c("darkblue", "#658c61") 
color_states <- c("#436340","#658c61","#86c97f")

# source in models 
source('code/Models.R') # source in C-R models
```

# Introduction

This file is for conceptually and quantitatively exploring the basis of resource competition theory (Tilman 1982, 1988). Some motivating questions include: 

a) From simple models of competition for resources, what qualititave system behaviors emerge? 
b) How do we understand community dynamics in an era of global change in light of resource competition theory? 

This document includes descriptions of models, equations, and assumptions proceeding from simple to more complex dynamics. Each model includes several plots to enable comparison across levels of increasing complexity. 

1. Diagrams/Equations (with verbal description)
2. Analytical Solution
3. Timeseries plots / Consumer/resource phase planes 
4. Bifurcation plot based on a parameter (list of possible dynamics)
5. Summary plot (model outcome heatmap w/two parameters)
 
# One consumer, one resource

## Model Description
A simple model for a single species of plant growing with one resource. This is a basic set of equations for modeling one consumer ($B$) and one resource ($R$). 

$$
\begin{aligned}
  \frac{dB}{dt} &= f \frac{gR}{1+ ghR}B - L B \\
  \frac{dR}{dt} &= r - \frac{gR}{1+ ghR}B \\
\end{aligned}
$$
Total biomass ($B$) is determined by the net balance between growth and loss rates. The growth rate per unit biomass is a function of resource availability, given by $G(R) = f \frac{g R}{1+ g h R}$, where $f$ is the resource conversion efficiency, $g$ is uptake rate constant, and $h$ controls the saturation rate of uptake. $G(R)$ is linear when $h = 0$, and saturates with increasing $R$ when $h > 0$. This model assumes a fixed loss rate $L$ per unit biomass. 

### Set Parameter Values
The following plots use the parameter values listed below. 
```{r}
# adjust parameter values 
p.list <- list(
  r=0.5, # supply rate of resource
  g=0.5, # growth rate
  f=1,   # conversion efficency of consumer
  h=0, # handling time
  L=0.7  # biomass loss rate
  )

p.table <- tibble(name = names(p.list), 
                  value = as.numeric(p.list),
                  description = c("resource supply rate", "per unit growth rate", "resource conversion efficiency", "uptake saturation", "biomass loss rate"))

knitr::kable(p.table)

```

## Analytical Solution 

For this system of equations, $R^*$ can be analytically calculated by setting $\frac{dB}{dt} = 0$. By this definition, $R^*$ is the resource level where biomass growth exactly balances loss and thus the minimum resource level required to sustain a non-zero biomass. 

```{r}

# Plot of isocline calculation
plot_R.star_calc <- with(as.list(p.list),{ # equation components as a function of resource level
  growth_curves <- tibble(R = seq(0,15, length.out=100), 
                          per_capita_growth = f*g*R/(1 + g*h*R), 
                          per_capita_mortality = rep(L, n=100))
  
  R_star <- L/(f*g-h*g*L)
  
  ggplot(growth_curves) + 
    geom_line(aes(x=R, y=per_capita_growth)) + 
    geom_line(aes(x=R, y=per_capita_mortality), linetype =2) + 
    geom_vline(aes(xintercept = R_star),linetype=3) + 
    labs(x = "R", y="G(R)", title = "Calculating R*")
}
)

plot_R.star_calc
```

Graphically, the value of $R^*$ is where the growth rate curve $G(R)$ curve crosses the loss-rate curve $L$. We see that increasing the loss rate (horizontal line) increases the value of $R^*$ given the monotonically increasing curve $G(R)$. 

```{r}
# Plot dependence of R* on parameter L
L.final <- 3
plot_R.star_L <- with(as.list(p.list),{
  L = seq(0,1.5, length.out = 1000)
  R_star <- L/(f*g-h*g*L)
  
  ggplot(tibble(L, R_star)) + 
    geom_line(aes(x=L, y=R_star)) +
    geom_vline(aes(xintercept = f/h), linetype=2) + # Loss rate where R* goes to infinity 
    labs(x = "Loss Rate (L)", y="R*", title = "R* dependence on L") 
}
)

plot_R.star_L + coord_cartesian(xlim = c(0, L.final))
```

The analytical solution of this simple system, $R^* = \frac{L}{fg - hgL}$ implies that $R^*$ increases with the loss rate ($L$). Note that $R^*$ (but not $B^*$) is independent of the resource supply rate ($r$). Introducing a handling time term $h > 0$ means that at a certain loss rate ($L = f/h$ = `r p.list$f/p.list$h`), the value of $R^*$ goes to infinity. 

```{r}
# Plot dependence of B* on parameter L 
eq.state <- with(as.list(p.list),{
  n = 1000
  r = seq(0,10, length.out = 10) #resource supply rate 
  L = seq(0,L.final, length.out = n) #loss rate
  p_eq <- expand.grid(r=r,L=L) %>% bind_cols(g = rep(g, n = n^2), f = rep(f, n = n^2), h = rep(h, n = n^2)) %>% 
    mutate(R_star = L/(f*g-h*g*L), B_star = r/(g*R_star))
  
  return(p_eq)
}) %>% pivot_longer(cols= R_star:B_star, names_to = "variable", values_to = "value")

plot_B.star_L <- ggplot(eq.state) + 
  geom_line(aes(x=L, y=value, linetype = variable, color = factor(r))) + 
  #scale_x_continuous(limits = c(0,10)) + 
  scale_y_continuous(limits = c(-5, 25)) + 
  scale_color_viridis_d(direction = -1, 
                        labels = function(x) round(as.numeric(x), 2),) + 
  labs(color = "Resoure Supply Rate (r)", 
       linetype = "Variable",
       x = "Loss Rate (L)", 
       y = "Equilibrium Value",
       title = "B* dependence on L")

plot_B.star_L
```

The amount of biomass $B^*$ that can be sustained depends on the resource supply rate (r). 

## Timeseries & Phase Plane
We can numerically integrate one instance. Analytical solutions predict values of `B` and `R` at equilbirium (dashed lines).
```{r}
# state variable values (initial values at beginning of "experiments")
i.state <- c(R=5, B=4) 

# parameter values 
# p.list <- list(
#   r=0.5, # supply rate of resource
#   g=0.5, # growth rate
#   f=1,   # conversion efficency of consumer 
#   h=0.5, # handling time 
#   L=1 # biomass loss rate
#   ) 

# default parameter values
# r <- 0.5 # supply rate of resource
# g <- 0.5 # growth rate
# f <- 1   # conversion efficiency of consumer
# h <- 0.5 # handling time
# L <- 0.7 # loss rate of consumer 

## adjust parameters
p.tseries <- p.list
# duration
Time <- 100
t.seq <- seq(0, Time, by = 0.01)
tseries <- ode(i.state,t.seq, one_sp_one_resource, p.tseries)

# analytical equilibrium resource levels 
state.eq <- with(as.list(p.tseries), {
  R.star <- L/(f*g-h*g*L)
  B.star <- r*f/L 
  
  return(list(R.star = R.star, B.star = B.star))
})
R.star <- state.eq$R.star
B.star <- state.eq$B.star

## plot dynamics around consumer and resource isoclines.

## time series
plot_temporal_dynamics <- ggplot(as.data.frame(tseries), aes(x = time)) +
  geom_line(aes(y = R, color = "Resource")) + 
  geom_line(aes(y = B, color = "Consumer Biomass")) +
  # analytical lines
  geom_hline(aes(yintercept = R.star, color = "Resource"), linetype = 2) + 
  geom_hline(aes(yintercept = B.star, color = "Consumer Biomass"), linetype = 2) + 
  ylab("Abundance") +
  xlab("Time") +
  scale_color_manual(values = color_palette, name = "") 

## Plot phaseplane 
# make arrow data
plot_arrows <- as.data.frame(cbind(tseries[-Time, ], tseries[-1, ]))
colnames(plot_arrows)[4:6] <- c("time.n1","R.n1", "B.n1")

# calculate isocline curves 

plot_isocline_dynamics <- ggplot() + 
  geom_segment(data = plot_arrows, aes(x = R, xend = R.n1, y = B, yend = B.n1), arrow = arrow(length = unit(0.1,"cm"))) +
  geom_point(aes(x = i.state[1], y = i.state[2]), shape = 21, fill = "white") +
  xlab("Resource") +
  ylab("Consumer Biomass") +  
  geom_hline(aes(yintercept = B.star, color = "B isocline"), linetype=2) + 
  geom_vline(aes(xintercept = R.star, color = "R isocline"), linetype=2) + 
  scale_color_manual(values = color_palette, name = "") 

# patchwork plots 
plot_temporal_dynamics / plot_isocline_dynamics
```

## Equilibrium Outcomes

### Parameter Scan 

Summary plot of the dynamics/equilibrium values of the model. 

```{r}
# Set-up storage matrices 
outlist <- list()
n = 100 # number of steps to break down the variables
out_end <- tibble(B_end = rep(NA, n), R_end = rep(NA,n)) # data frame of outputs

## Assign parameters in grid - choose two variables to scan over (r, L)
r = seq(0,5, length.out = n) #resource supply rate 
L = seq(0,1, length.out = n) #loss rate
plist <- expand.grid(r=r,L=L) %>% bind_cols(g = rep(1, n = n^2), f = rep(1, n = n^2), h = rep(1, n = n^2))

## Analytical solutions 
states.summary <- plist %>% 
  mutate(R.star = L/(f*g-h*g*L),
         B.star = r*f/L)

# Viz Heatmap of bimoass for r vs L 
ggplot(states.summary, aes(x=L, y=r)) + 
  geom_raster(aes(fill = B.star)) + 
  scale_fill_continuous(limits = c(0,20),) + 
  theme_minimal()

## Numerical solutions
# 
# #initial conditions 
# t = seq(from=0,to=100,by=0.5)
# B_0 = 5
# R_0 = 10
# y0 = c(B_0, R_0)
# 
# # Integrate & store output 
# for (i in 1:nrow(plist)) {
#   outlist[[i]] <- ode(y = y0, parms = plist[i,], func = one_sp_one_resource, times = t)
#   out_end[i, 1] <- outlist[[i]][,2] %>% tail(n=1)
#   out_end[i, 2] <- outlist[[i]][,3] %>% tail(n=1)
# }
# 
# out_full <- bind_cols(plist, out_end) %>% 
#   # add analytical solutions
#   mutate(R_star = L/(g*f)) %>% 
#   mutate(B_star = r/(g*R_star))

## Plot timeseries for a specific value
# m <- 50
# out_plot_p <- out_full[m,] # parameters
# out_plot <- as_tibble(outlist[[m]]) # timeseries
# colnames(out_plot) <- c("time", "B", "R")
# 
# # Viz 
# out_viz <- out_plot %>% pivot_longer(cols = B:R, names_to = "variable", values_to = "va1/lue")
# ggplot(out_viz) + geom_line(aes(x = time, y = value, color = variable)) + 
#   theme_bw() + 
#   geom_hline(aes(yintercept = out_plot_p$R_star, color = "R"), linetype = 2) + 
#   geom_hline(aes(yintercept = out_plot_p$B_star, color = "B"), linetype = 2)
```

Biomass heatmap shows that biomass at equilibrium increases with resource supply rate `r` and decreases with loss rate `L`. 
```{r}

# Line plot of B_end vs L 
states.summary %>% 
  filter(r %in% c(min(r), r[which.min(abs(r - median(r)))], max(r))) %>% # choose specific r values 
  ggplot(., aes(x=L, y = B.star)) + 
  geom_line(aes(color = factor(r))) + 
  scale_color_discrete(labels = function(x) round(as.numeric(x), 2)) + 
  labs(color = "Supply Rate (r)", 
       x = "Loss Rate (L)", 
       y = "Equilibrium Biomass")

```


# Two consumers, one resource

## Model Description
A simple model for two species ($B_1$ and $B_2$) competing for one resource ($R$). Consumers interact solely by competing for a shared resource. 

$$
\begin{aligned}
  \frac{dB_{1}}{dt} &= f_{1} \frac{g_{1}R}{1+ g_{1}h_{1}R}B_{1} - L_{1} B_{1} \\
  \frac{dB_{2}}{dt} &= f_{2} \frac{g_{2}R}{1+ g_{2}h_{2}R}B_{2} - L_{2} B_{2} \\
  \frac{dR}{dt} &= r - \frac{g_{1}R}{1+ g_{1}h_{1}R}B_{1} - \frac{g_{2}R}{1+ g_{2}h_{2}R}B_{2}\\
\end{aligned}
$$

Parameters are as described in the model for one consumer and one resource. 

### Set Parameter Values
The following plots use the parameter values listed below. 
```{r}
# adjust parameter values 
p.list <- list(
  r=0.5, # supply rate of resource
  
  ## species 1
  g1=1, # growth rate
  f1=0.5,   # conversion efficency of consumer
  h1=0.75, # handling time
  L1=0.5,  # biomass loss rate
  
  ## species 2
  g2=0.25, # growth rate
  f2=0.5,   # conversion efficency of consumer
  h2=0.5, # handling time
  L2=0.5  # biomass loss rate
  )

p.table <- tibble(name = names(p.list), 
                  value = as.numeric(p.list),
                  description = c("resource supply rate", 
                                  "sp. 1 per unit growth rate", 
                                  "sp. 1 resource conversion efficiency", 
                                  "sp. 1 uptake saturation", 
                                  "sp. 1 biomass loss rate",
                                  "sp. 2 per unit growth rate", 
                                  "sp. 2 resource conversion efficiency", 
                                  "sp. 2 uptake saturation", 
                                  "sp. 2 biomass loss rate"))

knitr::kable(p.table)

```


## Analytical Solution 

For this system of equations, $R^*$ can be analytically calculated by setting $\frac{dB}{dt} = 0$. By this definition, $R^*$ is the resource level where biomass growth exactly balances loss and thus the minimum resource level required to sustain a non-zero biomass. 

```{r}

# Plot of isocline calculation
with(as.list(p.list),{ # equation components as a function of resource level
  growth_curves <- tibble(R = seq(0,15, length.out=100), 
                          per_capita_growth_sp1 = f1*g1*R/(1 + g1*h1*R), 
                          per_capita_growth_sp2 = f2*g2*R/(1 + g2*h2*R), 
                          per_capita_mortality_sp1 = rep(L1, n=100),
                          per_capita_mortality_sp2 = rep(L2, n=100))  %>% pivot_longer(cols = 2:5, names_to = "name", values_to = "value") %>% 
  mutate(species = str_extract(name, "[a-z]{2}[0-9]"), # extract species 
         rate = str_split(name, pattern = "_", simplify = T)[,3]) # extract rate
  
  R.star.1 <- L1/(f1*g1 - g1*h1*L1)
  R.star.2 <- L2/(f2*g2 - g2*h2*L2)

  ggplot(growth_curves, aes(x=R, y = value, linetype = rate, color = species)) + 
  geom_line() + 
  geom_vline(aes(xintercept = R.star.1, color = "sp1"), linetype = 2) + # B1 R*
  geom_vline(aes(xintercept = R.star.2, color = "sp2"), linetype = 2) + # B2 R*
  labs(x = "R", y="Growth|Loss Rate ", title = "Calculating R*")
}
)
  
```

The species with the lower $R^*$ will outcompete the other. 

## Timeseries & Phase Plane
We can numerically integrate one instance. Analytical solutions predict values of $B_1$, $B_2$ and $R$ at equilbirium (dashed lines).
```{r}
# state variable values (initial values at beginning of "experiments")
i.state <- c(R=5, B1=4, B2=10) 

## adjust parameters
p.tseries <- p.list
# duration
Time <- 100
t.seq <- seq(0, Time, by = 0.01)
tseries <- ode(i.state,t.seq, two_sp_one_resource, p.tseries)

# analytical equilibrium resource levels 
state.eq <- with(as.list(p.tseries), {
  R.star.1 <- L1/(f1*g1 - g1*h1*L1)
  R.star.2 <- L2/(f2*g2 - g2*h2*L2)
  B1.star <- r*f1/L1 
  B2.star <- r*f2/L2
  
  return(list(R.star.1 = R.star.1, 
              R.star.2 = R.star.2, 
              B1.star = B1.star,
              B2.star = B2.star
              ))
})
R.star.1 <- state.eq$R.star.1
B1.star <- state.eq$B1.star

## plot dynamics around consumer and resource isoclines.
data.tseries <- as.data.frame(tseries) %>% pivot_longer(cols = c(R:B2), names_to = "variable", values_to = "value")

## time series
plot_temporal_dynamics <- ggplot(data.tseries, aes(x = time, y = value, color = variable)) +
  geom_line() + 
  # analytical lines
  geom_hline(aes(yintercept = R.star.1, color = "R"), linetype = 3) + 
  geom_hline(aes(yintercept = B1.star, color = "B1"), linetype = 2) + 
  geom_hline(aes(yintercept = state.eq$R.star.2, color = "R"), linetype = 4) + 
  geom_hline(aes(yintercept = state.eq$B2.star, color = "B2"), linetype = 2) + 
  ylab("Abundance") +
  xlab("Time")

plot_temporal_dynamics

## Plot phaseplane 
# make arrow data
plot_arrows <- as.data.frame(cbind(tseries[-Time, ], tseries[-1, ]))
colnames(plot_arrows)[5:8] <- c("time.n1","R.n1", "B1.n1", "B2.n1")

# calculate isocline curves 

plot_isocline_dynamics <- ggplot() + 
  geom_segment(data = plot_arrows, aes(x = R, xend = R.n1, y = B1, yend = B1.n1, color = "B1"), arrow = arrow(length = unit(0.1,"cm"))) +
  geom_segment(data = plot_arrows, aes(x = R, xend = R.n1, y = B2, yend = B2.n1, color = "B2"), arrow = arrow(length = unit(0.1,"cm"))) +
  geom_point(aes(x = i.state[1], y = i.state[2], color = "B1"), shape = 21, fill = "white") +
  geom_point(aes(x = i.state[1], y = i.state[3], color = "B2"), shape = 21, fill = "white") +
  xlab("Resource") +
  ylab("Consumer Biomass") +  
  geom_hline(aes(yintercept = state.eq$B1.star, color = "B1"), linetype=2) + 
  geom_hline(aes(yintercept = state.eq$B2.star, color = "B2"), linetype=2) + 
  geom_vline(aes(xintercept = state.eq$R.star.1, color = "B1"), linetype=3) +
  geom_vline(aes(xintercept = state.eq$R.star.2, color = "B2"), linetype=4) 


# patchwork plots 
plot_temporal_dynamics / plot_isocline_dynamics
```

## Equilibrium Outcomes 

There are a number of possible outcomes at equilibrium. 

  1. No species can be sustained. 
  2. R* (sp1) < R* (sp2) - Sp. 1 wins
  3. R* (sp2) < R* (sp1) - Sp. 2 wins

### Parameter Scan 
```{r}
# Set-up storage matrices 
outlist <- list()
n = 100 # number of steps to break down the variables
out_end <- tibble(B_end = rep(NA, n), R_end = rep(NA,n)) # data frame of outputs

## Assign parameters in grid - choose two variables to scan over (r, L)
p1 = seq(0.1,5, length.out = n) #parameter 1
p2 = seq(0.1,1, length.out = n) #parameter 2
p.dataframe <- bind_rows(p.list) %>%
  slice(rep(row_number(), n^2)) %>% 
  select(-L1, -g2) #select which parameters to replace 
p.grid <- expand.grid(L1 = p1,
                      g2 = p2) %>% 
  bind_cols(p.dataframe)

## Analytical solutions 
states.summary <- p.grid %>% 
  mutate(R.star.1 = L1/(f1*g1 - g1*h1*L1),
         R.star.2 = L2/(f2*g2 - g2*h2*L2),
         B1.star = r*f1/L1,
         B2.star = r*f2/L2) %>% 
  mutate(B.star = if_else(R.star.1 < R.star.2, B1.star, B2.star)) %>% 
  mutate(sp = if_else(R.star.1 < R.star.2, 1, 2)) %>%
  mutate(sp = as.factor(sp)) 

head(states.summary)
# Viz Heatmap of bimoass for r vs L 
ggplot(states.summary, aes(x=L1, y=g2)) + 
  geom_raster(aes(fill = sp, alpha = B.star)) 
#theme_minimal()
```


# Two consumers, one fluctuating resource 

## Model Description
A simple model for two species ($B_1$ and $B_2$) competing for one resource ($R$) with a supply rate that varies in time. Consumers interact solely by competing for a shared resource. 

$$
\begin{aligned}
  \frac{dB_{1}}{dt} &= f_{1} \frac{g_{1}R}{1+ g_{1}h_{1}R}B_{1} - L_{1} B_{1} \\
  \frac{dB_{2}}{dt} &= f_{2} \frac{g_{2}R}{1+ g_{2}h_{2}R}B_{2} - L_{2} B_{2} \\
  \frac{dR}{dt} &= r(t) - \frac{g_{1}R}{1+ g_{1}h_{1}R}B_{1} - \frac{g_{2}R}{1+ g_{2}h_{2}R}B_{2}\\
  r(t) &= \sin(\beta t)
\end{aligned}
$$
## Possible extensions

- Time varying $r$. 
- Decomposition of biomass supplies $r$ 
- Different shapes to growth curves (exponential)

# SCRAP CODE
```{r, include = F, eval=FALSE}

## create resource isocline
R.iso <- expression(L/(f*g-h*g*L)) # set R = 0 and solved algebraically
R.star <- eval(R.iso)
B.iso <- expression(r/(g*L/(f*g-h*g*L)))
B.star <- eval(B.iso)

```

