---
title: "Makes shrub paper figures"
author: "Ruby An"
date: "2024-07-13"
output:
  pdf_document: default
  html_document:
    number_sections: true
editor_options:
  chunk_output_type: console
---

# Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 1, digits = 2) #set to two decimal 

## Packages
library(tidyverse)
library(deSolve)
library(knitr)
library(ggpubr)
library(cowplot)

## Plot Theme
theme_set(theme_classic())
```

# Functions
```{r}
# Model functions ------

make_parameters <- function(heights, # heights of S species
                            # set species generic parameters
                            a = 10,
                            r = 2, 
                            b = 0.6,
                            beta = 5,
                            m = 0.1, 
                            k = 2) {
  # b : biomass density
  # beta : power law allometry
  # m : mortality
  # a : photosynthetic rate
  # r : respiration rate
  # k : light capture per ramet
  # c : crown area per ramet
  # L_above : light at top of canopy
  
  # of species S
  S <- length(heights)
  
  # demographic parameters
  p_phys <- setNames(list(a, r, b, beta, m, k), 
                   c("a","r","b","beta", "m","k"))
  
  # calculate minimum light requirement
  biomass <- b*heights^beta
  fi <- a/biomass
  mi <- r/biomass + m
  ki <- rep(k,S)
  ui = mi/(fi*k)

  # Set parameter list for numerical solution to Equilibrium
  p <- setNames(list(fi,mi,ki,1), c("u", "f","m","k","L_above"))
  
  return(setNames(list(ui,p), c("ui", "p")))
}

simulate_ramets <- function(t,y,p) {
  
  # number of species
  S <- length(y) 
  
  # p is a list of species-specific parameters: 
  # f : fecundity
  # m : mortality
  # k : light capture per ramet
  # L_above : light at top of canopy
  
  dy.dt <- rep(NA, S)
  with(as.list(p), {
    #first species 
    dy.dt[1] = L_above*f[1]*(1-exp(-k[1]*y[1]))-m[1]*y[1] 
    
    if (S > 1) {
      for (i in 2:S) {
        dy.dt[i] = L_above*f[i]*exp(-k[i]*sum(y[1:(i-1)]))*(1-exp(-k[i]*y[i])) - m[i]*y[i]
      }
      return(list(c(dy.dt)))
    } else { # only one species
      return(list(c(dy.dt)))
    }
  })
}

simulate_ramets_physio <- function(t,y,p) {
  
  # number of species
  S <- length(y) 
  
  # p is a list of species-specific parameters: 
  # h : height of ramet
  # b : biomass density
  # beta : power law allometry
  # m : mortality
  # a : photosynthetic rate
  # r : respiration rate
  # k : light capture per ramet
  # c : crown area per ramet
  # L_above : light at top of canopy
  
  dy.dt <- rep(NA, S)
  with(as.list(p), {
    #first species 
    dy.dt[1] = (1/(b*h[1]^beta))*((a*L_above)*(1-exp(-k[1]*c[1]*y[1]))/y[1] - r)*y[1] - m[1]*y[1] 
    
    if (S > 1) {
      for (i in 2:S) {
        dy.dt[i] = (1/(b*h[i]^beta))*((a*L_above*e^(-k[i]*c[i]*sum(y[1:(i-1)])))*(1-exp(-k[i]*c[i]*y[i]))/y[i] - r)*y[i] - m[1]*y[1] 
      }
      return(list(c(dy.dt)))
    } else { # only one species
      return(list(c(dy.dt)))
    }
  })
}

calculate_ramet_eq <- function(u,p) {
  
  # u is the light requirement per species that satisfies the invasion condition
  
  with(as.list(p), {
    # set vectors
    S = length(u) # f, m, and k should be the same dimension, number of species
    y = rep(NA, S)
    L = rep(NA, S) # light level
    
    # First species 
    L_total = 1 # light level at top of canopy
    
    # Search interval, with total light
    lower_bound <- (1/k[1])*log(L_total/u[1]) # first derivative of f_x
    upper_bound <- L_total/(k[1]*u[1]) 
    
    # numerical root find implicit solution for eq 
    num_solve <- uniroot(function(x) L_total*(1-exp(-k[1]*x)) - k[1]*u[1]*x, 
                         lower = lower_bound,
                         upper = upper_bound)
    
    # First Species EQ density
    y[1] <- max(0,num_solve$root)
    
    # Light environment for shorter species
    L[1] <- L_total*exp(-k[1]*y[1])
    
    # Species 2:S
    
    if(S > 1) {
      for (i in 2:S) {
        
        # define search interval
        lower_bound <- (1/k[i])*log(L[i-1]/u[i]) # first derivative of f_x
        upper_bound <- L[i-1]/(k[i]*u[i])
        
        # numerical root find
        num_solve <- uniroot(function(x) L[i-1]*(1-exp(-k[i]*x)) - k[i]*u[i]*x, 
                             lower = lower_bound,
                             upper = upper_bound)
        y[i] <- max(0, num_solve$root)
        
        # set light below canopy i
        L[i] <- L[i-1]*exp(-k[i]*y[i])
      }
    L_above <- c(L_total, L[1:(i-1)])
    }  
    else { # do nothing, only one specie
      L_above <- L_total
    }
    
    env_eq <- tibble(L_above = L_above, u = u, y = y, L_below = L) %>% 
      # set extinction. floor
      # mutate(y = round(y, 4)) %>% 
      # calculate change in light level below canopy
      mutate(dL = -(L - lag(L_above,default = L_total))) %>% 
      mutate(species = paste0("y",1:S)) %>% 
      
      # clean up
      rename(y_eq=y) %>% 
        
      # feasiblity
      mutate(feasible = if_else(round(y_eq,digits = 3) == 0, FALSE, TRUE)) %>% 
      mutate(light_acquired = L_above-u) %>% 
      
      # check time to reproduction
      mutate(light_per_ramet = L_above/y_eq)
    
    return(env_eq)
  })
}
```


# Figure 1 - Model Cartoon 


# Figure 2 - Multi-Species model 

## Parameters from "Process Informed Metric"
This code chooses a set number "S" of species randomly along the "light requirement (u_i)" trait axes. It that back-calculates the heights and biomass of each species from the light requirement (u). From this, and the parameters conserved across species, you can calculate fecundity and mortality to pass to the ODE solver. 

```{r}
# Parameters ---- 
# Choose number of species + initial conditions
S <- 30

# Set generic parameters
a <- 10
r <- 2
b <- 0.6
beta <- 5
m <- 0.1 # mortality
k <- 2 # crown area + light extinction

# Generate species PIM
u_min <- r/(a*k)
u_min
u <- sort(runif(S, min = u_min, max = 1), decreasing =T) # u_min = 0.01

# Ramet morphology & demography
height = ((a*k*u-r)/(m*b))^(1/beta)
biomass <- b*height^beta
fi <- a/biomass
mi <- r/biomass + m
ki <- rep(k,S)
ui = mi/(fi*k)

p <- setNames(list(fi,mi,ki,1), c("f","m","k","L_above"))
```


### Equilibrium solution
```{r}
# Equilibrium
eq_table <- calculate_ramet_eq(u,p) %>% mutate(species = paste0("y",1:S)) %>% 
  rename(y_eq=y) %>% 
  
  # check time to reproduction
  mutate(light_per_ramet = L_above/y_eq)%>% 
  mutate(height = ((a*k*u-r)/(m*b))^(1/beta)) %>% 
  mutate(biomass = b*height^beta) %>% 
  mutate(fecundity = a/biomass) %>% 
  mutate(mortality = r/biomass + m) %>% 
  
  # feasiblity
  mutate(feasible = if_else(y_eq == 0, FALSE, TRUE)) %>% 
  mutate(light_acquired = L_above-u)

y_max <- max(0.5,max(eq_table$y_eq*1.05))

# Assess Equilibrium Abundance
model_eq_analytical <- eq_table %>% 
  filter(feasible == TRUE) %>% 
  ggplot(aes(x=u, y=y_eq)) + 
    #geom_rect(aes(xmin = L_below, xmax = u,ymin = 0, ymax = max(eq_table$y_eq), alpha = 0.9)) + #alpha = if_else(y_eq==0, Inf, y_eq))) +
    scale_color_manual(values = c("red", "black")) +
    geom_hline(yintercept=0, color = "red", linetype = 2) + 
    #geom_vline(xintercept = eq_table$L_below, color = "grey") + 
  geom_point(data = eq_table, aes(x = u, y = y_eq, color = feasible), size=3) + 
  labs(x = "Light Requirement", y = "Ramet Density") + 
  lims(y = c(0,y_max))

# # Normalized Equilbirium Abundance
# eq_table %>%
#   mutate(y_norm = y_eq/light_acquired) %>%
#   mutate(feasible = if_else(y_eq == 0, FALSE, TRUE)) %>%
#   ggplot(aes(x=u, y=y_norm))+
#     scale_color_manual(values = c("red", "black")) +
#     geom_hline(yintercept=0, color = "red", linetype = 2) +
#     #geom_vline(xintercept = eq_table$L_below, color = "grey") +
#     #geom_rect(aes(xmin = L_below, xmax = u,ymin = 0, ymax = max(eq_table$y_eq), alpha = if_else(y_eq==0, Inf, y_eq))) +
#   geom_point(aes(color = feasible, x = u, y = y_norm), size = 3) +
#   geom_function(fun = function(x) 1/(k*x), color = "grey")

```

### ODE Simulation
```{r}
# Simulation
# Set timespan
t = seq(from=0,to=1/p$m[1]*2000,by=0.1)
#t = seq(from=0,to=400,by=0.01)

# Initial conditions
y0 = rep(0.01,S)
#y0 = eq_table$y_eq

# Integrate
out = ode(y=y0,times=t,func=simulate_ramets,parms=p);

# Arrange Data
data <- as_tibble(as.data.frame(out))
colnames(data) <- c("time", paste0("y",1:S))
viz_data <- data %>% pivot_longer(cols = contains("y"), 
                                  names_to = "species", values_to = "value") %>% 
  left_join(eq_table)

# Plot 
model_eq_sim <- ggplot(viz_data, aes(x = time, y = value)) + 
  geom_line(aes(color = species, group = species))+ 
  geom_hline(aes(yintercept = y_eq, color = species), linetype = 3)+
  theme_bw() + 
  labs(x = "Time", y="Ramet Density") + 
  scale_color_discrete(name = "Species") + 
  lims(y = c(0,y_max))

```

## Parameters from Physiology
This code reads the parameter values written in "ramet_parameters.csv" that describe parameters consistent across species. It generates a list of species heights, then calculates the minimum light requirement required per species (key trait that determines coexistence potential). 

It then generates a set of derived parameters (fecundity, mortality) for each species that can be passed to the ODE solver. 

```{r}
# generate parameters for S species from physiology
S <- 30

# species generic parameters
a <- 10
r <- 2
b <- 0.6
beta <- 5
m <- 0.1 # mortality
k <- 2 # crown area + light extinction

p_phys <- setNames(list(a, r, b, beta, m, k), 
                   c("a","r","b","beta", "m","k"))
  # p_phys is a list of species-specific parameters: 
  # b : biomass density
  # beta : power law allometry
  # m : mortality
  # a : photosynthetic rate
  # r : respiration rate
  # k : light capture per ramet
  # c : crown area per ramet
  # L_above : light at top of canopy

# calculate minimum u "minimum light requirement"
u_min <- r/(a*k)

# calculate maximum height (for minimum light requirement = 1)
h_max <- ((a*k - r)/(m*b))^(1/beta)

# generate uniform distribution of heights
heights <- sort(runif(S, min = 0, max = h_max), decreasing =T)

# calculate minimum light requirement
biomass <- b*heights^beta
fi <- a/biomass
mi <- r/biomass + m
ki <- rep(k,S)
ui = mi/(fi*k)

# Set parameter list for numerical solution to Equilibrium
p <- setNames(list(fi,mi,ki,1), c("f","m","k","L_above"))
```

### Analytical EQ
```{r}
# Calculate Equilibria
(eq_table <- calculate_ramet_eq(ui,p))

# Assess Equilibrium Abundance
y_max <- max(0.5,max(eq_table$y_eq*1.05)) # y limit for graph

(model_eq_analytical <- eq_table %>% 
  filter(feasible == TRUE) %>% 
  ggplot(aes(x=u, y=y_eq)) + 
    #geom_rect(aes(xmin = L_below, xmax = u,ymin = 0, ymax = max(eq_table$y_eq), alpha = 0.9)) + #alpha = if_else(y_eq==0, Inf, y_eq))) +
    scale_color_manual(values = c("red", "black")) +
    geom_hline(yintercept=0, color = "red", linetype = 2) + 
    #geom_vline(xintercept = eq_table$L_below, color = "grey") + 
  geom_point(data = eq_table, aes(x = u, y = y_eq, color = feasible), size=3) + 
  scale_alpha_manual(values = c(1,0))+
  labs(x = "Light Requirement", y = "Ramet Density") + 
  lims(y = c(0,y_max)))

# Light Availabililty

canopy_above <- c(0, cumsum(eq_table$y_eq))
light_below <- c(1, eq_table$L_below)
light_env <- tibble(canopy_above = canopy_above, light_below=light_below)

light_env %>% 
  ggplot(aes(x=1-light_below, y = canopy_above)) + 
  geom_point() + 
  scale_y_reverse()
```

### ODE Simulation
```{r}
# Simulation
# Set timespan
t = seq(from=0,to=1/p$m[1]*100,by=0.01)
#t = seq(from=0,to=1000,by=0.01)

# Initial conditions
y0 = rep(0.1,S)
#y0 = eq_table$y_eq/3

# Integrate
out = ode(y=y0,times=t,func=simulate_ramets,parms=p);

# Arrange Data
data <- as_tibble(as.data.frame(out))
colnames(data) <- c("time", paste0("y",1:S))
viz_data <- data %>% pivot_longer(cols = contains("y"), 
                                  names_to = "species", values_to = "value") %>% 
  left_join(eq_table)

# Plot 
model_eq_sim <- ggplot(viz_data, aes(x = time, y = value)) + 
  geom_line(aes(color = species, group = species))+ 
  geom_hline(aes(yintercept = y_eq, color = species), linetype = 3)+
  theme_bw() + 
  labs(x = "Time", y="Ramet Density") + 
  scale_color_discrete(name = "Species") + 
  lims(y = c(0,y_max))
```

### Abundance vs Height distribution 

#### 30 species
```{r}

# number of iterations 
iter <- 10^4

# number of species 
S <- 30

# species generic parameters
a <- 10
r <- 2
b <- 0.6
beta <- 5
m <- 0.1 # mortality
k <- 2 # crown area + light extinction

# maximum height 
h_max <- ((a*k - r)/(m*b))^(1/beta)
h_higher <- h_max*1.25

# make a dataframe to fill
df <- tibble(iter=numeric(),
                 species=character(), 
                 height=numeric(), 
                 u=numeric(),
                 y_eq=numeric()) 

# calculate eq. abundance for multiple iterations
for (i in seq(1:iter)){

    # choose heights (uniform distribution)
  heights <- sort(runif(S, min = 0, max = h_higher), decreasing =T)
  
  # calculate minimum light requirement + species parameters 
  biomass <- b*heights^beta
  fi <- a/biomass
  mi <- r/biomass + m
  ki <- rep(k,S)
  ui = mi/(fi*k) # min light requirement

  # Set parameter list for numerical solution to equilibrium
  p <- setNames(list(fi,mi,ki,1), c("f","m","k","L_above"))
  
  # calculate anayltical eq abundance
  eq_abundance <- calculate_ramet_eq(ui, p) %>% 
    mutate(iter = i) %>% 
    rename(species = species)
  eq_abundance$height <- heights
  
  # bind to dataframe
  return_eq <- eq_abundance %>% 
    select(iter, species, height, u, y_eq, everything())
  
  df <- bind_rows(df, return_eq)
}


# height bins 
nbins = 1000
H <- seq(0,h_higher,length.out = nbins+1)
U <- seq(0,1,length.out = nbins+1)

N = N_u = L_a = L_b <- numeric(nbins) # initialize vectors

for (i  in seq(1:nbins)) {
  
  # bin by height or light requirement (u)
  N[i]=mean(df$y_eq[df$height >= H[i] & df$height < H[i+1]])
  N_u[i]=mean(df$y_eq[df$u >= U[i] & df$u < U[i+1]])
  
  # light above and below
  L_a[i]=mean(df$L_above[df$height >= H[i] & df$height < H[i+1]])
  L_b[i]=mean(df$L_below[df$height >= H[i] & df$height < H[i+1]])
}

# plot abundance vs height
plot_height_dist <- plot(H[1:nbins],N)

# plot abundance vs light requirement
plot(U[1:nbins],N_u)

# plot light vs height
ggplot(df, aes(x=height, y = L_above)) + 
  geom_point()
plot(H[1:nbins],L_a)

plot(H[1:nbins],L_a-L_b)
plot(L_a, N)

```

#### 100 Species
Tried the code but it got stuck finding the analytical solution... upper and lower bound were reveresed or something like that. 

### Figure 2 
```{r}
(fig2 <- plot_grid(model_eq_analytical, model_eq_sim, plot_height_dist))
```

# Figure 3 - Multiple Mechanisms for Productivity

1. Choose the species heights from the previous simulation. 
2. Recalculate the light requirement with a higher photosynthetic rate.
3. Recalculate species abundances, creating df_a.


## Increase a
```{r}

a_multiplier = 1.5

# make a dataframe to fill
df_a <- tibble(iter=numeric(),
                 species=character(), 
                 height=numeric(), 
                 u=numeric(),
                 y_eq=numeric()) 

# calculate eq. abundance for multiple iterations
for (i in seq(1:iter)){

    # choose heights (uniform distribution)
  heights <- df$height[df$iter == i]
  
  # calculate minimum light requirement + species parameters 
  biomass <- b*heights^beta
  fi <- a*a_multiplier/biomass
  mi <- r/biomass + m
  ki <- rep(k,S)
  ui = mi/(fi*k) # min light requirement

  # Set parameter list for numerical solution to equilibrium
  p <- setNames(list(fi,mi,ki,1), c("f","m","k","L_above"))
  
  # calculate anayltical eq abundance
  eq_abundance <- calculate_ramet_eq(ui, p) %>% 
    mutate(iter = i) 
    
  eq_abundance$height <- heights
  
  # bind to dataframe
  return_eq <- eq_abundance %>% 
    select(iter, species, height, u, y_eq, everything())
  
  df_a <- bind_rows(df_a, return_eq)
}

# height bins 
nbins = 1000
H <- seq(0,h_higher,length.out = nbins+1)
U <- seq(0,1,length.out = nbins+1)

N = N_u = L_a = L_b <- numeric(nbins) # initialize vectors

for (i  in seq(1:nbins)) {
  
  # bin by height or light requirement (u)
  N[i]=mean(df_a$y_eq[df_a$height >= H[i] & df_a$height < H[i+1]])
  N_u[i]=mean(df_a$y_eq[df$u >= U[i] & df_a$u < U[i+1]])
  
  # light above and below
  L_a[i]=mean(df_a$L_above[df$height >= H[i] & df_a$height < H[i+1]])
  L_b[i]=mean(df_a$L_below[df$height >= H[i] & df_a$height < H[i+1]])
}

# plot abundance vs height
plot_height_dist <- plot(H[1:nbins],N)

# plot abundance vs light requirement
plot(U[1:nbins],N_u)

# plot light vs height
ggplot(df, aes(x=height, y = L_above)) + 
  geom_point()
plot(H[1:nbins],L_a)

plot(H[1:nbins],L_a-L_b)
plot(L_a, N)

df_a <- df_a %>% mutate(sim = "a_inc")


## comparison 
df_comp <- df %>% mutate(sim = "normal") %>% 
  bind_rows(df_a) 

df_binned <- df_comp %>% 
  mutate(bin = round(height/h_higher,digits = 3))

df_binned %>% 
  group_by(bin,sim) %>% 
  summarize(mean_abundance = mean(y_eq)) %>% 
  ggplot(aes(x=bin*h_higher, y = mean_abundance, color = sim)) + 
  geom_point(size=3) + 
  labs(x = "Height", y = "Abundance")

## Species richness
df_comp %>% 
  group_by(iter,feasible,sim) %>% 
  summarize(richness=n()) %>% 
  ggplot(aes(x = sim, y = richness)) + 
  geom_boxplot(aes(fill = feasible))

```

### Increase a infilling only
```{r}
# make a dataframe to fill
df_infill <- tibble(iter=numeric(),
                 species=character(), 
                 height=numeric(), 
                 u=numeric(),
                 y_eq=numeric()) 

# calculate eq. abundance for multiple iterations
for (i in seq(1:iter)){

    # choose heights (uniform distribution)
  heights <- df$height[df$iter == i]
  
  # calculate minimum light requirement + species parameters 
  biomass <- b*heights^beta
  fi <- a*a_multiplier/biomass
  mi <- r/biomass + m
  ki <- rep(k,S)
  ui = mi/(fi*k) # min light requirement

  # Set parameter list for numerical solution to equilibrium
  p <- setNames(list(fi,mi,ki,1), c("f","m","k","L_above"))
  
  # calculate anayltical eq abundance
  eq_abundance <- calculate_ramet_eq(ui, p) %>% 
    mutate(iter = i) %>% 
  eq_abundance$height <- heights
  
  # bind to dataframe
  return_eq <- eq_abundance %>% 
    select(iter, species, height, u, y_eq, everything())
  
  df_infill <- bind_rows(df_infill, return_eq)
}

```

## Increase k

## Increase d

# Figure 3 - Case #1 Infilling
This case describes a scenario where all S increase in productivity (a = photosynthetic rate) the same amount. 

## Figure 3
```{r}
S = 30
set.seed(1)

# Normal Tundra 
p <- make_parameters(S=S, a=10)
ui_norm = p$m/(p$f*p$k)

norm_eq <- calculate_ramet_eq(ui_norm,p) %>%
  # scenario
  mutate(scenario = "normal")

# Warmer Tundra: increase productivity (a) of all species
p_w <- p
p_w$f <- p$f*1.5
ui_warm <-p_w$m/(p_w$f*p_w$k)

warm_eq <- calculate_ramet_eq(ui_warm,p) %>% 
  # scenario
  mutate(scenario = "warmer")

compare_eq <- norm_eq %>% bind_rows(warm_eq)

y_max = max(compare_eq$y_eq*1.01)

## VIZ

plot_compare_eq <- compare_eq %>% ggplot(aes(x=u, y=y_eq)) + 
    geom_hline(yintercept=0, color = "red", linetype = 2) + 
    geom_point(aes(x = u, y = y_eq, color = species, shape=feasible), size=3) + 
  scale_shape_manual(values = c(1,16)) +
  labs(x = "Light Requirement", y = "Ramet Density") + 
  lims(y=c(0,y_max)) + 
  facet_wrap(vars(scenario)) + 
  theme_bw()

(fig3 <- plot_compare_eq)

```

# Figure 4 - Case #2 Range Expansion
Add a taller species 

```{r}
# Double productivity (a) 
p_w <- p
p_w$f <- p$f*2
ui_warm <-p_w$m/(p_w$f*p_w$k)

# Add taller species
u_tree <- runif(1, min = max(ui_warm), max = 1)
#u_tree <- 0.8
ui_range <- c(u_tree, ui_warm)

# parameters
height = ((a*k*u_tree-r)/(m*b))^(1/beta)
biomass <- b*height^beta
f_range <- c(a/biomass, p_w$f)
m_range <- c(r/biomass, p_w$m)

p_range <- setNames(list(f_range,m_range,k= rep(k,length(m_range)),1), c("f","m","k","L_above"))

# EQ conditions w/range expansion

compare_range_eq <- calculate_ramet_eq(ui_range, p_range) %>% mutate(species = paste0("y",0:S)) %>%
  # scenario
  mutate(scenario = "range_expansion") %>% 
  
  #add to other scenario
  bind_rows(warm_eq)

## Plot 
plot_range_eq <- compare_range_eq %>% ggplot(aes(x=u, y=y_eq)) + 
    geom_hline(yintercept=0, color = "red", linetype = 2) + 
    geom_point(aes(x = u, y = y_eq, color = species, shape=feasible), size=3) + 
  scale_shape_manual(values = c(1,16)) +
  labs(x = "Light Requirement", y = "Ramet Density") + 
  lims(y=c(0,y_max)) + 
  facet_wrap(vars(scenario)) + 
  theme_bw()

(fig4 <- plot_range_eq)

```

# Figure 5 - Case #3 Height Plasticity 

```{r}

```


#### Extra Code

##### Old "Parameters from Physiology" at set sequence of intervals code
```{r}

# read species-agnostic parameters
# param_file <- "3_ramet_light_model/ramet_parameters.csv"
# p_table <- read_csv(param_file)
# p_list <- setNames(as.list(p_table$value), p_table$parameter) 

# calculate species-specific parameters
min_height <- 2
max_height <- 120 
sp_heights <- tibble(species = paste0("y", seq(1:S)), value = seq(max_height, min_height, length.out = S)) %>%  # indexed tallest to shortest 
  mutate(parameter = "h", description = "height", unit = "cm")

# calculate species-specific rates: f, m, kc 
## f : fecundity
## m : mortality
## k : light capture per ramet, k is the extinction coefficient and c is canopy leaf area


p_sp <- with(as.list(p_list), { # p is a list of species-specific parameters: 
  k <- rep(k, S) # light capture, for future crown size differences
  fi <- a/(b*sp_heights$value^beta) # fecundity
  mi <- r/(b*sp_heights$value^beta) + m
  
  p <- setNames(list(fi,mi,k), c("f", "m", "k"))
  return(p)
})

# set environmental parameters
## L_above : light available above the canopy 
env_values <- p_table%>% filter(species == "env") %>% pull(value)
env_names <- p_table %>% filter(species == "env") %>% pull(parameter)
p_env <- setNames(as.list(env_values), env_names) 

# full parameter list: species + environment parameters 
p <- c(p_sp, p_env)
(sp_table <- sp_heights %>% 
  rename(height = value) %>% 
  select(-parameter,-description, -unit) %>% 
  mutate(biomass = p_list$b*height^p_list$beta) %>% 
  mutate(f = p$f, m = p$m, k = p$k) %>% 
  mutate(ui = m/(f*k))) # calculate light requirement (u) per species

ui <- sp_table$ui
#histogram of light requirement
hist(ui[ui<1],breaks=100)
```

