---
title: "Shrub Transplant Experimental Design"
author: "Ruby An"
date: "8/20/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("randomizr")
library("psych")
library("tidyverse")
```

## Experimental Design

I have 10 blocks, each with 8 treatments, for a total of 80 plots. 

Step 1: Generate a unique ID for every plot. 
Step 2: Make a list of the treatments. 

```{r design}
blocks <- c(1:10)
plots <- c(1:8)
plot_id <- c(1:80)

treatments <- c("C", "W", "N", "S", "W+N", "W+S", "N+S", "W+N+S")

set.seed(1)

sample(treatments)


sheet 

## Testing the randomizr package
blocks <- rep(c("A", "B","C"), times = c(50, 100, 200))
Z <- block_ra(blocks = blocks)
table(blocks, Z)

## Testing the psych package

create.randomization.plan <- function(n.per.group, factors, seed=NULL){
  factor.lengths <- sapply(factors,length)
  if(!is.null(seed)){
    set.seed(seed)
  }
  plan <- as.data.frame(block.random(n.per.group*prod(factor.lengths),
                                     factor.lengths))
  for(i in 1:length(factors)){
    plan[,colnames(plan)==names(factors)[i]] <-
      factor(plan[,colnames(plan)==names(factors)[i]])
    levels(plan[,colnames(plan)==names(factors)[i]]) <- factors[[i]]
  }
  return(plan)
}

factors <- list(fertilizer=c("N", "_"),
                snow=c("S", "_"),
                warming=c("W", "_"))
 
plan <- create.randomization.plan(10, factors, seed = 284) %>%
  
  # add helpful columns 
  mutate(treatment = str_c(fertilizer,snow,warming, sep = "_")) %>% bind_cols(plot_id=plot_id) %>% 
  select(plot_id, blocks, everything())

headTail(plan, top=16)


write_csv(plan, "shrub_expmt_design.csv")
```


## Blocking the Snow 


I have 10 blocks, each with 8 treatments, for a total of 80 plots. 

Step 1: Generate a unique ID for every plot. 
Step 2: Make a list of the treatments. 

```{r design}
blocks <- c(1:10)
plots <- c(1:8)
plot_id <- c(1:80)

treatments <- c("C", "W", "N", "S", "W+N", "W+S", "N+S", "W+N+S")

## Testing the psych package

create.randomization.plan <- function(n.per.group, factors, seed=NULL){
  factor.lengths <- sapply(factors,length)
  if(!is.null(seed)){
    set.seed(seed)
  }
  plan <- as.data.frame(block.random(n.per.group*prod(factor.lengths),
                                     factor.lengths))
  for(i in 1:length(factors)){
    plan[,colnames(plan)==names(factors)[i]] <-
      factor(plan[,colnames(plan)==names(factors)[i]])
    levels(plan[,colnames(plan)==names(factors)[i]]) <- factors[[i]]
  }
  return(plan)
}

factors <- list(fertilizer=c("N", "_"),
                warming=c("W", "_"))
 
plan <- create.randomization.plan(20, factors, seed = 284) %>% 
  rename(sub_blocks = blocks)

headTail(plan, top=5)

## create snow blocks

snow_factors <- list(snow = c("S", "_"))
snow_plan <- create.randomization.plan(10, snow_factors, seed = 284) %>% 
  bind_cols(sub_blocks = seq(1:20))
headTail(snow_plan, top=16)


master_plan <- left_join(plan, snow_plan) %>% 

  # add helpful columns m
  mutate(treatment = str_c(fertilizer, snow, warming, sep = "_")) %>% 
  mutate(plot_num = rep(seq(1:8), 10)) %>% 
  mutate(plot_id = seq(1:80)) %>% 
  select(blocks, plot_num, treatment, everything())

headTail(master_plan, top = 16)

write_csv(master_plan, "shrub_expmt_design_snow_blocked.csv")
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
