---
title: "simple_light_competition"
author: "Ruby An"
date: "2024-03-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 1, digits = 2) #set to two decimal 

## Packages
library(tidyverse)
library(deSolve)
library(knitr)
library(ggpubr)

## Plot Theme
theme_set(theme_classic())
```

## Model Description 

This is a simplified version of the tundra light competition model. 

## Functions

You can also embed plots, for example:

```{r functions}

eq_light_extinction <- function(u,p) {
  
  with(as.list(p), {
     # set vectors
  S = length(u) # f, m, and k should be the same dimension, number of species
  y = rep(NA, S)
  L = rep(NA, S)
  
  # First species 
  L_above = 1 # light level at top of canopy
  
  # Search interval, with total light
  lower_bound <- (1/k[1])*log(L_above/u[1]) # first derivative of f_x
  upper_bound <- L_above/(k[1]*u[1]) 
  
  # numerical root find implicit solution for eq 
  num_solve <- uniroot(function(x) L_above*(1-exp(-k[1]*x)) - k[1]*u[1]*x, 
                       lower = lower_bound,
                       upper = upper_bound)
  
  # First Species EQ density
  y[1] <- max(0,num_solve$root)
  
  # Light environment for shorter species
  L[1] <- exp(-k[1]*y[1])
  
  # Species 2:S
  
  if(S > 1) {
    for (i in 2:S) {
      
        # define search interval
        lower_bound <- (1/k[i])*log(L[i-1]/u[i]) # first derivative of f_x
        upper_bound <- L[i-1]/(k[i]*u[i])
        
        # numerical root find
        num_solve <- uniroot(function(x) L[i-1]*(1-exp(-k[i]*x)) - k[i]*u[i]*x, 
                             lower = lower_bound,
                             upper = upper_bound)
        y[i] <- max(0, num_solve$root)
        
        # set light below canopy i
        L[i] <- L[i-1]*exp(-k[i]*y[i])
      }
  }  else { # do nothing, only one specie
      }
  
  env_eq <- tibble(L_below = L, u = u, y = y) %>% 
    mutate(dL = -(L - lag(L,default = 1)))
  
  return(env_eq)
  })
}

simulate_light_extinction <- function(t,y,p) {
  
  # number of species
  S <- length(y) 
  
  # p is a list of parameters: 
  # f
  # m
  # k
  
  dy.dt <- rep(NA, S)
  
  #first species 
  dy.dt[1] = f[1]*(1-exp(-k[1]*y[1]))-m[1]*y[1] 
  
  if (S > 1) {
    for (i in 2:S) {
      dy.dt[i] = f[i]*exp(-k[i]*sum(y[1:(i-1)]))*(1-exp(-k[i]*y[i])) - m[i]*y[i]
    }
    return(list(c(dy.dt)))
  } else { # only one species
    return(list(c(dy.dt)))
  }
}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Equilibrium Solutions - Analytical

## Simulations 

### Parameters
```{r}
# Choose number of species + initial conditions
S <- 1
y0 <- rep(0.01, S)

# Set parameters
f <- sort(runif(S)*10)
m <- rep(1,S)
k <- rep(1,S)
p <- setNames(list(f,m,k), c("f","m","k"))
```

### Analytical Solutions
```{r}
# Analytical solutions
u <- with(as.list(p), { 
  u <- m/(f*k)
  return(u)
})

(eq_density <- eq_light_extinction(u,p))

tibble(species = paste0("y",1:S), u = u)

```

### Integration
```{r}
# Set timespan
t = seq(from=0,to=100,by=0.01)

# Integrate
out = ode(y=y0,times=t,func=simulate_light_extinction,parms=p);

# Arrange Data
data <- as_tibble(as.data.frame(out))
colnames(data) <- c("time", paste0("y",1:S))
viz_data <- data %>% pivot_longer(cols = contains("y"), names_to = "variable", values_to = "value")

# Plot 
ggplot(viz_data, aes(x = time, y = value)) + 
  geom_line(aes(color =variable)) + 
  geom_hline(yintercept = eq_density$y, color = eq_density$u, linetype = 2) + 
  theme_bw() 
```


## Eigenvalues 



